---
title: "Insurance_Fraud_Detection"
output: html_document
date: "2024-04-06"
---
Data source : 

# Predicting Complaint Types in Insurance Companies

Predicting complaint types in insurance companies can provide several benefits and avenues for increasing productivity and efficiency:

1. **Improved Customer Service**: 
   - By accurately predicting complaint types, insurance companies can proactively address customer issues and concerns. 
   - This can lead to higher customer satisfaction, improved retention rates, and positive word-of-mouth referrals.

2. **Resource Allocation**: 
   - Understanding the types and frequency of complaints allows companies to allocate resources more effectively. 
   - They can prioritize areas that receive the most complaints, ensuring that resources are allocated where they are most needed.

3. **Risk Management**: 
   - Analyzing complaint types can help identify potential risks and areas of vulnerability within the company's operations. 
   - By addressing these issues proactively, companies can mitigate risks and avoid costly legal disputes or regulatory penalties.

4. **Product Development**: 
   - Insights from complaint data can inform product development and innovation. 
   - By understanding customer pain points and preferences, companies can develop new products or improve existing ones to better meet customer needs.

5. **Process Optimization**: 
   - Analyzing complaint data can highlight inefficiencies or bottlenecks in internal processes. 
   - Companies can use this information to streamline workflows, automate repetitive tasks, and improve overall operational efficiency.

6. **Training and Education**: 
   - Identifying common complaint types can guide training programs for employees. 
   - Companies can provide targeted training to address specific areas of concern, improving employee competency and performance.

7. **Regulatory Compliance**: 
   - Compliance with regulatory requirements is essential in the insurance industry. 
   - Analyzing complaint data can help ensure that companies are meeting regulatory standards and requirements, reducing the risk of non-compliance and associated penalties.

8. **Competitive Advantage**: 
   - Companies that effectively analyze and respond to complaint data can gain a competitive advantage in the market. 
   - By providing superior customer service and addressing customer concerns promptly, companies can differentiate themselves from competitors and attract more customers.

```{r}

insurance <- read.csv("Insurance_complaints__All_data.csv")

dim(insurance)

```
```{r}

column_counts_df <- data.frame(Column_Name = character(), Unique_Count = numeric(), stringsAsFactors = FALSE)

# Iterate over columns of the original data frame
for (col in names(insurance)) {
  # Calculate unique count of values in each column
  unique_count <- length(unique(insurance[[col]]))
  # Append column name and unique count to the new data frame
  column_counts_df <- rbind(column_counts_df, data.frame(Column_Name = col, Unique_Count = unique_count))
}

# Display the new data frame
print(column_counts_df)

```
```{r}

selected_columns <- c("Complaint.type", "Complaint.filed.against", "Reason.complaint.filed",
                      "Coverage.type", "Coverage.level", "Others.involved", "Keywords")
selected_df <- insurance[selected_columns]

column_counts_df <- data.frame(Column_Name = character(), Unique_Count = numeric(), stringsAsFactors = FALSE)

# Iterate over columns of the original data frame
for (col in names(selected_df)) {
  # Calculate unique count of values in each column
  unique_count <- length(unique(selected_df[[col]]))
  # Append column name and unique count to the new data frame
  column_counts_df <- rbind(column_counts_df, data.frame(Column_Name = col, Unique_Count = unique_count))
}

# Display the new data frame
print(column_counts_df)

```


```{r}

library(ggplot2)
library(janitor)

insurance_df <- clean_names(insurance)

# Frequency table for 'Complaint_Type'
complaint_type_freq <- table(insurance$complaint_type)
print(complaint_type_freq)
```

```{r}

theme_set(theme_bw())

# Bar plot for 'Complaint_Type' with counts displayed on top of each bar
ggplot(data = insurance_df, aes(x = complaint_type)) +
  geom_bar(fill = "skyblue", color = "black") +
  geom_text(stat = 'count', aes(label = ..count..), vjust = -0.2) +  # Add counts on top of each bar
  labs(title = "Distribution of Complaint Types") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))


# Bar plot for 'Coverage_Type'
ggplot(data = insurance_df, aes(x = coverage_type)) +
  geom_bar(fill = "lightgreen", color = "black") +
  geom_text(stat = 'count', aes(label = ..count..), vjust = -0.2) +
  labs(title = "Distribution of Coverage Types") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```


```{r}

library(dplyr)
library(ggplot2)

insurance_df |>
  group_by(coverage_level)|>
  summarise(counts = n())|>
  filter(counts > 1000)|>
  ggplot(aes(x = coverage_level, y = counts)) +
  geom_bar(stat = "identity", fill = "skyblue", color = "black") +
  labs(title = "Distribution of Coverage Levels (Counts > 1000)") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

```


```{r}

ggplot(data = insurance_df, aes(x = complaint_filed_by)) +
  geom_bar(fill = "skyblue", color = "black") +
  geom_text(stat = 'count', aes(label = ..count..), vjust = -0.2) +  # Add counts on top of each bar
  labs(title = "Distribution of Complaint Types") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

```

```{r}

insurance_df |>
  group_by(complaint_filed_by)|>
  summarise(counts = n())|>
  filter(counts > 1000)|>
  ggplot(aes(x = complaint_filed_by, y = counts)) +
  geom_bar(stat = "identity", fill = "skyblue", color = "black") +
  labs(title = "Distribution of Coverage Levels (Counts > 1000)") +
  geom_text(aes(label = counts), vjust = -0.3, color = "black", size = 3) + 
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

```

```{r}

ggplot(data = insurance_df, aes(x = confirmed_complaint)) +
  geom_bar(fill = "skyblue", color = "black") +
  geom_text(stat = 'count', aes(label = ..count..), vjust = -0.2) +  # Add counts on top of each bar
  labs(title = "Distribution of Complaint Types") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

```

```{r}

library(arules)

# Convert dataframe to transactions
transactions <- as(insurance_df, "transactions")

# Perform association rule mining
rules <- apriori(transactions)
summary(rules)
# Assuming 'rules' is your association rules object
inspect(rules)


```

```{r}

# Set minimum thresholds
min_support <- 0.1
min_confidence <- 0.8
min_lift <- 1.5

# Filter rules based on thresholds
selected_rules <- rules[quality(rules)$support > min_support &
                        quality(rules)$confidence > min_confidence &
                        quality(rules)$lift > min_lift]

#head(inspect(selected_rules))

```

```{r}

# Load required packages
library(ggplot2)
library(arulesViz)

# Convert rules to data frame
rules_df <- as(selected_rules, "data.frame")

# Plot scatter plots for support vs. confidence and support vs. lift
plot(rules_df, method = "scatterplot")

```

```{r}

# Set minimum thresholds
min_support <- 0.10
min_confidence <- 0.80
min_lift <- 5

# Filter rules based on thresholds
selected_rules <- rules[quality(rules)$support > min_support &
                        quality(rules)$confidence > min_confidence &
                        quality(rules)$lift > min_lift]

head(inspect(selected_rules))

```


```{r}

predictive_variables <-  c( "complaint_filed_against", "complaint_filed_by", "reason_complaint_filed", "confirmed_complaint", 
            "received_date", "closed_date",  "coverage_type", "coverage_level", "others_involved",
            "respondent_role", "respondent_type", "complainant_type", "keywords")

response_variables <- c("complaint_type")

# duration needs to calculated in between "received_date" and "closed_date"
#keywords needs to be imputed 

```
# Models for Handling Categorical Data

## Decision Trees
Decision trees are versatile models that can handle both categorical and numerical variables. They partition the feature space into mutually exclusive regions based on the predictor variables and make predictions by following the decision path from the root to the leaf nodes.

## Random Forests
Random forests are an ensemble learning method that combines multiple decision trees to improve prediction accuracy. They work well with categorical variables and are robust to overfitting.

## Gradient Boosting Machines (GBM)
GBM is another ensemble learning technique that builds a series of decision trees sequentially, where each tree corrects the errors of the previous one. GBM can handle categorical variables and often yields high predictive performance.

## Logistic Regression
Logistic regression is a classical statistical model used for binary classification tasks. It can be extended to handle multi-class classification problems using techniques like one-vs-rest or multinomial logistic regression.

## Support Vector Machines (SVM)
SVMs are powerful classifiers that can handle both linear and non-linear decision boundaries. They can be applied to categorical data by using appropriate kernel functions, such as the polynomial kernel or radial basis function (RBF) kernel.

## Naive Bayes
Naive Bayes classifiers are probabilistic models based on Bayes' theorem with the assumption of independence between predictors. They are particularly well-suited for text classification tasks but can also be applied to categorical data with many levels.

## Neural Networks
Deep learning models, such as multilayer perceptrons (MLPs) or convolutional neural networks (CNNs), can handle high-dimensional categorical data. However, they often require a large amount of data and computational resources for training.

## Ensemble Methods
Apart from random forests and gradient boosting machines, other ensemble methods such as AdaBoost and XGBoost can also be effective for handling categorical data with many levels.


